---
title: "Unit 1 Paper Appendix, Version 2"
author: "Xiangsen Yang"
date: "1/17/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


## obtaining the data and package

```{r cars}
library(confintr)
library(Rmisc)

data = read.delim("https://dept.stat.lsa.umich.edu/~bbh/s485/data/israel_severe_covid_2021_08_15.tsv")
list(data)
```

## Overview

The computations in this document explore the vaccinated and unvaccinated hospitalizations rate from different age groups in Israel, which includes the calculation for Agresti-Coull confidence intervals for different binomial variables.

## CI of the Agresti-Coull interval

Combining BCD equation (5) on p. 108 with notation defined in section 3.1.2 on p.108, the Agresti-Coull interval is:

$CI_{AC}=\tilde{p}\pm\kappa\sqrt{\tilde{p}(1-\tilde{p})}\sqrt{\frac{1}{n+\kappa^2}}$,

Denote $\tilde{X}=X+\kappa^2/2$, $\tilde{n}=n+\kappa^2$. 


# 1.

AC interval might includes negative number, but the rate is on the interval [0,1], so we limit our Agresti-Coull interval on[0,1] 

**_Result of interval in the unvaccinated :_**
```{r }

#interval for the underlying rates of severe disease in the unvaccinated
Interval_U <- as.data.frame(matrix(nrow=0,ncol=3)) 
names(Interval_U)<-c("ages","PU_interval_2.5","PU_interval_97.5")
a <- 1
while (a <= length(data[,1])){
  R_Agresti <- ci_proportion(data[a,'cases_unvax'],n =data[a,'pop_unvax'],
                             type = "Agresti-Coull" )
  
  Interval_U <- rbind(Interval_U, 
                      data.frame(ages = data[a,'ages'] ,
                                 PU_interval_2.5 = round(R_Agresti$interval[1], digits = 7),
                                 PU_interval_97.5=round(R_Agresti$interval[2], digits = 7)))
  a <- a + 1
}
Interval_U
```
**_Result of interval in the vaccinated :_**

```{r }
#interval for the underlying rates of severe disease in the vaccinated
Interval_V <- as.data.frame(matrix(nrow=0,ncol=3)) 
names(Interval_V)<-c("ages","PV_interval_2.5","PV_interval_97.5")
b <- 1
while (b <= length(data[,1])){
  R_Agresti <- ci_proportion(data[b,'cases_vax'],
                             n =data[b,'pop_vax'],type = "Agresti-Coull" )
  Interval_V <- rbind(Interval_V, 
                      data.frame(ages = data[b,'ages'] ,
                                 PV_interval_2.5 = round(R_Agresti$interval[1], digits = 7),
                                 PV_interval_97.5 = round(R_Agresti$interval[2], digits = 7)))
  b <- b + 1
}
Interval_V
```
The overall result, as shown below:
```{r}
PV_PU <- merge(Interval_U,Interval_V,by="ages")
PV_PU
```
| Ages   | Lowest Case Rate for Unvaccinated  | Highest Case Rate for Unvaccinated  | Lowest Case Rate for Vaccinated  | Highest Case Rate for Vaccinated |
|--------|:---------------:|---------------:|--------------:|--------------:|
| 12-15 | 0.0000000 | 0.0000163 | 0.0000000 | 0.0000251 |
| 16-19 | 0.0000003 | 0.0000611 | 0.0000000 | 0.0000108 |
| 20-29 | 0.0000043 | 0.0000402 | 0.0000000 | 0.0000047 |
| 30-39 | 0.0000340 | 0.0001093 | 0.0000000 | 0.0000081 |
| 40-49 | 0.0001097 | 0.0002470 | 0.0000048 | 0.0000188 |
| 50-59 | 0.0002859 | 0.0005638 | 0.0000192 | 0.0000448 |
| 60-69 | 0.0005797 | 0.0010127 | 0.0000672 | 0.0001128 |
| 70-79 | 0.0013841 | 0.0026051 | 0.0001614 | 0.0002432 |
| 80-89 | 0.0017751 | 0.0035723	| 0.0003932 | 0.0005825 |
| 90+   | 0.0030765 | 0.0083531 | 0.0002398 | 0.0006151 |
# 2. a point estimate and corresponding confidence interval for the vaccine efficacy

**_Result of interval for the vaccine efficacy :_**
```{r }
#Efficiency Mean Point Estimation
c <-1
while (c <= length(PV_PU[,1])) {
  point_U<- mean(PV_PU$PU_interval_2.5[c],PV_PU$PU_interval_97.5[c])
  point_V<- mean(PV_PU$PV_interval_2.5[c],PV_PU$PV_interval_97.5[c])
  PV_PU[c,'Point_est'] <- round(1-point_V/point_U,3)
  c <- c+1
}
PV_PU[which(is.nan(PV_PU$Point_est)),'Point_est']<-1

#Confidence Intervals for Vaccine Efficiency
d <-1
while (d <= length(PV_PU[,1])) {
  #Calculate the vaccine efficiency according to the rate of the first question
  efficacy_Lower <- 1-PV_PU$PV_interval_2.5[d]/PV_PU$PU_interval_2.5[d]
  efficacy_Upper <- 1-PV_PU$PV_interval_97.5[d]/PV_PU$PU_interval_97.5[d]
  #confidence interval
  Ci_P <- CI(c(efficacy_Lower,efficacy_Upper))
  
  if(is.nan(Ci_P[1])){
    Ci_P[1]<-1
  } 
  #The vaccine efficiency cannot be negative, and the negative value is set to 0
  if(is.nan(Ci_P[3]) || Ci_P[3]<0 ){
    Ci_P[3]<-0
  } 
  PV_PU$efficacy_Lower[d] <- efficacy_Lower
  PV_PU$efficacy_Upper[d] <- efficacy_Upper
    
  PV_PU$Eff_Lower[d] <- round(Ci_P[3],3)
  PV_PU$Eff_Upper[d] <- round(Ci_P[1],3)

  d <- d+1
}

PV_PU[,c('Point_est','Eff_Lower','Eff_Upper')]


```
| Ages   | estimated vaccine efficacy   | Lowest vaccine efficacy  | Highest vaccine efficacy  | 
|--------|:---------------:|---------------:|---------------:|
| 12-15 | 1.000 | 0.000 | 1.000 | 
| 16-19 | 1.000 | 0.000 | 2.035 | 
| 20-29 | 1.000 | 0.199 | 1.684 | 
| 30-39 | 1.000 | 0.492 | 1.434 | 
| 40-49 | 0.956 | 0.734 | 1.146 | 
| 50-59 | 0.933 | 0.849 | 1.005 | 
| 60-69 | 0.884 | 0.858 | 0.915 | 
| 70-79 | 0.883 | 0.747 | 1.043 | 
| 80-89 | 0.778 | 0.436	| 1.179 | 
| 90+   | 0.922 | 0.897 | 0.952 |

# 3. The estimated vaccine efficacy as calculated without adjustment for age, (Where pv and pu) denote severe COVID-19 rates across vaccinated and unvaccinated Israelis aged 12 and older.

```{r }
#PV and PU for all age groups
PU <- ci_proportion(sum(data[,'cases_unvax']),n =sum(data[,'pop_unvax']),type = "Agresti-Coull" )
PV <- ci_proportion(sum(data[,'cases_vax']),n =sum(data[,'pop_vax']),type = "Agresti-Coull" )

#Vaccine Efficiency at All Ages
efficacy_U_2.5 <- 1-PV$interval[1]/PU$interval[1]
efficacy_U_97.5 <- 1-PV$interval[2]/PU$interval[2]
all <- c(efficacy_U_2.5,efficacy_U_97.5)
#confidence interval
CI(all)
```
The mean estimated vaccine efficacy is about 68%.
